---
title: "Statistical assignment 4"
author: "Simone Long_135288"
date: "29/02/2020"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

```{r}

setwd("C:/Users/simon/OneDrive/Documents/datan3_2019")

library(tidyverse)
Data8 <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
setwd("C:/Users/simon/OneDrive/Documents/datan3_2019")
Stable <- read_tsv("data/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )
    ```

1. Univariate distribution (20 points).

    ```{r}
 
library(ggplot2)
Data %>%    
    ggplot(aes(h_payn_dv)) + 
    geom_freqpoly() +
    xlab("Net Monthly Pay") +
    ylab("Number of respondents")

    ```

The median monthly pay of UK residents (among those who responded) appears to fall around £1500. Using this sample to generalise, we can assume that the majority of UK residents fall in a middle/working class income bracket.

2. Line chart (20 points).

    ```{r}

Data %>%
    ggplot(aes(x = h_age_dv, y = h_payn_dv, group=sex_dv)) +
    geom_smooth(aes(linetype=sex_dv), color="black") +
    xlim(15, 65) +
    xlab("Age")+
    ylab("Monthly Earnings") +
    labs(linetype="Sex")
        
    ```

Among this sample's respondents, it seems that men tend to earn more than women. This disparity is particularly present between those aged 30 and 50 (with a dramatic difference in monthly earnings), as competition for high-earning jobs tends to be fiercest among those in this age bracket.

3. Faceted bar chart (20 points).

    ```{r}
#DONE
Data <- Data %>%
    group_by(sex_dv) %>%   
    mutate(medpay = median(h_payn_dv, na.rm = TRUE))

Data %>%
drop_na(placeBorn, sex_dv) %>%
ggplot(aes(x = sex_dv, y = medpay)) +
        geom_bar(stat = 'identity') +
        facet_wrap( ~ placeBorn) +
        ylim(0, 2000) +
        xlab("Sex") +
        ylab("Median monthly net pay")
    
    ```

There is no significant difference in the gender pay gap across different countries of origin. Men have a consistently higher median net monthly pay than women, highlighting the global epidemic of sexism in the working world.

4. Heat map (20 points).

    ```{r}
#ARGH    
 Data2 <- Data %>% 
    group_by(h_gor_dv, placeBorn) %>%
    mutate(meanage = mean(h_age_dv))
    
Data2 %>%
    drop_na(h_gor_dv, placeBorn) %>%
    ggplot(aes(x = h_gor_dv, y = placeBorn, fill = meanage)) +
    geom_tile(stat = 'identity') +
    theme(axis.text.x = element_text(angle = 90)) +
    xlab("Region") +
    ylab("Country of Birth") +
    labs(fill="Mean Age")
    
    ```

The average age of the population tends not to be significantly young or old relative to country of origin and region settled in in the UK. But it should be noted that the average age of those from Nigeria has a younger trend (i.e. below the age of 40), especially among those who have settled in Scotland, South West England and Yorkshire, with the oldest average coming from Jamaica and settling in Scotland.

5. Population pyramid (20 points).

    ```{r}
    
    Data %>%
    ggplot(mapping = aes(x = h_age_dv, fill = sex_dv)) +
        geom_bar(data=subset(Data, sex_dv == "female"))+
        geom_bar(data=subset(Data, sex_dv == "male"), aes(y=..count..*(-1)))+
         coord_flip() +
        labs(fill="Sex") +
        xlab("Age") +
        ylab("n")
        
    
    ```
    
The age of the sample is fairly evenly distributed among men and women. There is, however, a marginally greater number of women sampled, especially those age 30-50. 